The Efficient Handling of Guards \\in the Design of RPython's Tracing JIT

### 2012-09-07 talk/vmil2012/paper.tex
Add some notes about the changes to the tables

Figure~\ref{fig:benchmarks} summarizes the total number of operations that were
recorded during tracing for each of the benchmarks and what percentage of these
operations are guards. The static number of operations was counted on the unoptimized
and optimized traces. The figure also shows the overall optimization rate for
operations, which is between 69.4\% and 83.89\%, of the traced operations and the
optimization rate of guards, which is between 65.8\% and 86.2\% of the
operations. This indicates that the optimizer can remove
most of the guards, but after the optimization pass these still account for
15.2\% to 20.2\% of the operations being compiled and later executed.
The frequency of guard operations makes it important to store the associated
information efficiently and also to make sure that guard checks are executed
quickly.

\subsection{Guard Failures}
\label{sub:guard_failure}
The last point in this discussion is the frequency of guard failures.
Figure~\ref{fig:failing_guards} presents for each benchmark a list of the
relative amounts of guards that ever fail and of guards that fail often enough that a bridge is compiled.\footnote{
    The threshold used is 200 failures. This rather high threshold was picked experimentally to give
    good results for long-running programs.
}

The numbers presented for guards that have a bridge represent the
failures up to the compilation of the bridge and all executions of the then
attached bridge.

@@

From Figure~\ref{fig:failing_guards} we can see that only a very small amount
of all the guards in the compiled traces ever fail. This amount varies between
2.4\% and 5.7\% of all guards. As can be expected, even fewer, only 1.2\% to 3.6\% of all guards fail often
enough that a bridge is compiled for them.
Also, of all failing guards a few fail extremely often
and most fail rarely. Reinforcing this notion the figure shows that, depending on the
benchmark, between 0.008\% and 0.225\% of the guards are responsible for 50\%
of the total guards failures.
These results emphasize that as most of the guards never
fail it is important to make sure that the successful execution of a guard does
not have unnecessary overhead.

@@

%\todo{remove this section and the figures}
%\begin{figure*}
%    \include{figures/ops_count_table}
%    \caption{Relative numbers of operations in the traces generated for
%    different benchmarks}
%    \label{fig:ops_count}
%\end{figure*}
%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{figures/ops_pie.pdf}
%\caption{Relative frequency of operations before and after optimization}
%\label{fig:ops_pie}
%\end{figure*}
\bibliographystyle{abbrv}
\bibliography{zotero,paper}
\listoftodos
\end{document}

>>>>

Figure~\ref{fig:benchmarks} summarizes\footnote{In all tables the minimum and
maximum values for each column are highlighted in dark/light gray.} the total number of operations that were
recorded during tracing for each of the benchmarks and what percentage of these
operations are guards. The static number of operations was counted on the unoptimized
and optimized traces. The figure also shows the overall optimization rate for
operations, which is between 69.4\% and 83.89\%, of the traced operations and the
optimization rate of guards, which is between 65.8\% and 86.2\% of the
operations. This indicates that the optimizer can remove
most of the guards, but after the optimization pass these still account for
15.2\% to 20.2\% of the operations being compiled and later executed.
The frequency of guard operations makes it important to store the associated
information efficiently and also to make sure that guard checks are executed
quickly.

\subsection{Guard Failures}
\label{sub:guard_failure}
The last point in this discussion is the frequency of guard failures.
Figure~\ref{fig:failing_guards} presents for each benchmark a list of the
relative amounts of guards that ever fail and of guards that fail often enough that a bridge is compiled.\footnote{
    The threshold used is 200 failures. This rather high threshold was picked experimentally to give
    good results for long-running programs.
}
It also contains sparklines depicting the failure rates for the failing guards
in decreasing order, each normalized to the most failing guard.
The numbers presented for guards that have a bridge represent the
failures up to the compilation of the bridge and all executions of the then
attached bridge.

@@

From Figure~\ref{fig:failing_guards} we can see that only a very small amount
of all the guards in the compiled traces ever fail. This amount varies between
2.4\% and 5.7\% of all guards. As can be expected, even fewer, only 1.2\% to 3.6\% of all guards fail often
enough that a bridge is compiled for them.
Also, of all failing guards a few fail extremely often
and most fail rarely. Reinforcing this notion the figure shows that, depending on the
benchmark, between 0.008\% and 0.225\% of the guards are responsible for 50\%
of the total guards failures. Even considering 99.9\% of guard failures the
relative amount of guards does not rise above 3\%.
The colored dots in the sparklines correspond to 50\%, 99\% and 99.9\%.
These results emphasize that as most of the guards never
fail it is important to make sure that the successful execution of a guard does
not have unnecessary overhead.

@@

%\todo{remove this section and the figures}
%\begin{figure*}
%    \include{figures/ops_count_table}
%    \caption{Relative numbers of operations in the traces generated for
%    different benchmarks}
%    \label{fig:ops_count}
%\end{figure*}
%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{figures/ops_pie.pdf}
%\caption{Relative frequency of operations before and after optimization}
%\label{fig:ops_pie}
%\end{figure*}
\bibliographystyle{abbrv}
\bibliography{zotero,paper}
%\listoftodos
\end{document}
